---
title: "Week_11_Stats_Notes"
format: html
---

# review

Rides(exp)~i~ = 1071 + 177Celc -402Wkn + 18.5Celc\*Wkn \# make sub i throughout

Rides(exp) = alpha + beta1Celc + beta2Wkn + beta3Celc\*Wkn

what does the model think the weekday/weekend difference is? it depends! (on the temp)

regression numbers actually mean shit so... know what it is

# longitudinal data

2 things you can do

-   model trajectory

-   make better causal inferences

```{r}
library(tidyverse)
library(panelr)
```

```{r}
data("WageData") # long form data
```

## panel possibilities

-   X doesn't change (like sex)

    -   mixed models and growth curves (random slopes on time)

-   X changes once ("treatment" introduction)

    -   pre/post

    -   dif-in-dif

-   X changes at different times, in the same direction (having a child)

    -   two-way fixed effects

    -   dif-in-dif for staggered treatments

-   X changes at different times, in any direction (reversible treatments like exercising)

    -   two-way fixed effects

    -   mixed models

    -   within-between models/correlated random effects

-   any of the above + X is determined dynamically

## types of variance

-   within-subject (aka within-group)

-   between-subject (aka between-group)

ICC works the same. but level 2 variance tends to be greater for panel data bc it is a function of the basis of similarity

```{r}
line_plot(WageData, lwage, id = "id", wave = "t", subset.ids = TRUE, n.random.subset = 10)
```

## types of variables

-   time-constant

    -   only vary between units (like race, birth year)

-   time-varying

    -   measured over time but almost always have with and between variance (earnings, satisfaction)

```{r}
d <- panel_data(WageData,
                id = id,
                wave = t)
are_varying(d)
```

fem (gender), ed (education), and blk (race) only have between-person variance

\^ have ICC of 100%

## when x doesn't change

regular mixed model just with people now.

but there is no time in the model! assumes that fluctuations/errors across time are random

simplest fix is just adding time as a normal additive variable in the model (but this still assumes that time effects everyone in the same way)

### 4 ways to model time

-   time affects everyone the same

-   time affects everyone in the same group the same (college degree vs not, for ex.)

    -   group interaction on time with no random slope on time

-   every person gets their own time effect (latent growth curve model)

    -   random slope on time with no group interaction on time

-   combo of 2 and 3

    -   random slope and group interaction on time

## multilevel models DO NOT improve causal inference

all it does is add another error term that is (like the og error term) assumed to be independent of the predictors

it is **NOT** another control variable. it does nothing for the DAG

if you want better causal inference from panel data, you have to use within-person variance or time-varying treatments

we should compare people to themselves bc a lot less shit changes for a person than it might between people

## when x changes once

### pre-post

"fixed-effects"

we want to compare y1t1 - y0t0 with y0t1 - y0t0

Yt1 = betaXt1 + mu + error1

Yt2 = betaXt2 + mu + error2

mu are things that don't change but affect y (this is the fixed effect)

subtract the equations and get (Yt2-Yt1) = beta + (mu - mu) + (e2 - e1)

\^ there are too many variables, so we might have to assume that e2 and e1 are equal, which is crazy, but this means beta is just the difference between Y at the two times

to deal with potential e1/e2 differences, we gotta compare within the person to between the people (aka do did-in-dif)

### dif-in-dif

difference within person in difference between people

compare just world people with the treatment and world people to get "just treatment"

we can solve everything with 4 numbers

aka the difference of the time betas between treated people over time and untreated people over time

**this does not require the treated/untreated groups to be similar! just that we can assume time affects individual people in both groups in the same way**

E(y) = beta0 + beta1treat + b2post +beta3treat\*post **(with multilevel model clustered at person level)**

**what we want for the ATT is beta3**

and all the covariates/possible confounders are in there already!! yipppeeee!!!!!!!!!!!!

\^ adding adjustment variables will NOT change beta3 (but it might change the beta on treatment)

#### DiD assumptions

-   parallel trends: in the absence of the treatment, the slopes would be the same

-   no anticipation: treated group doesn't change their behavior in anticipation of the treatment (knowing about the treatment doesn't change shit at time 0)

we can "test" these if we have 2 (or more) pre-treatment observations

-   parallel trends: if the 2 groups are running parallel before the treatment then the suggestion is that the effect of time is the same

    -   try a dif-in-dif between the first 2 (non-treated) observations, and if there is an observable effect, then the parallel trends assumption is violated

#### DiD augmentations

aka adding matching or weighting on pre-treatment covariates and then doing DiD

yay doubly robust estimates! (but the standard errors are fucked up again unless you do a million bootstraps)

## when x changes at different times (irreversibly)

differential timing/staggered treatments

### two-way fixed effects

(secret heads-up that COVID made TWFE a whole thing and maybe it doesn't work the way we want bc this assumes that the treatment effect is the same no matter how long you've had it)

trying to remove 2 major sources of confounding

-   individual unit fixed effect

-   effect of being observed at a particular time

we don't give a shit about the people that are untreated (or treated) the whole time (okay we give a little shit bc they help tell us about time)

#### linear probability models (aka ols on a dummy/binary variable)

the output is in **percentage points** NOT percentage change

poor? = betaMother~ij~ + mu~i~ + theta~j~ + error~ij~

mu is all the shit about a person that doesn't change but affects the chance of being poor

theta is the stuff about time periods that makes poverty more or less common/likely

this can make predictions that someone's probability is less than 0 or more than 100

#### conditional logit models (also for binary outcome)

the same equation as above but this time in log odds

this lets us condition out the number of times someone is in poverty to predict why those numbers are where they are instead of in a different configuration

however we can't use the people that are always in the same outcome status (like always in poverty) booooo

so basically we like the LPM more

#### cans and can'ts of TWFE

can

-   adjust for ALL time-constant confounders

-   adjust for effects of specific time periods

-   identify homogeneous treatment effects **IF** assumptions are met

    -   assumption is that the effect is constant forever and not diminishing or aggregating

can't

-   rule out time-varying confounders (without adding them to the model)

-   rule out reverse causality

    -   can find things wiggling together really good, but can't know which wiggler is the source

-   deal with heterogeneous treatment effects

#### adding adjustment variables

you gotta manually throw in time-varying confounders

**but you always gotta think about the DAG**

## when x changes at different times (and you can go back and forth)

still TWFE

basically just look at if the wiggles follow each other

**we can only estimate the effect for the kinds of people that we see switching**

(unless we make some assumptions) this is ATT vs conditional ATT

this also works for a continuous x

**all TWFE are variance weighted** (the people with the most variance contribute the most to the estimate)

# mixed-models for time-varying treatments

puttin' it all together

its basically just 2 different regressions (within and between) sharing coefficients

but you are back to the assumptions about no correlation between x and the 2 error terms

this lets us use all the people (even the ones who stayed at the same x the whole time)

you can actually just put two betas in there
