---
title: "Stats_Homework_04"
format: html
---

# Part One

### Question 1

A\) The error is 1

B\) The residual is 0.9

### Question 2

Top left: Y = X + D

Top right: Y = X + D (using D as a proxy for A)

Bottom left: Y = X

Bottom right: Y = X

### Question 3

A\) 76.185

B\) 19.693

C\) 306.553

D\) 778.965

E\) Yes, the coefficient is significant. The estimates are 12.13 and 13.03 standard deviations away from 0 (respectively).

### Question 4

A\) 110.230 - 3.162Edu

B\) Annual hours worked would increase by 59.638

C\) The relationship gets less positive as education increases.

D\) Adding too many polynomial terms makes the relationship between the variables increasingly difficult to interpret.

### Question 5

A\) The coefficient suggests that being a homeowner is associated with a 50.174 hour increase in the number of hours worked compared to someone who is not a homeowner. However, this estimate is not statistically significant.

B\) The model suggests that people with 4 children work 150.492 less hours than people with 3 children.

C). We cannot tell from this table. We would need to rerun the regression using "2 children under 5" as the base of comparison rather than "0 children under 5".

### Question 6

A\) A one unit increase in education is associated with either a 56.079 or a 110.073 increase in hours worked depending on whether or not the individual is a homeowner.

B\) The coefficient means that being a homeowner causes each one-unit increase in education to cause an increase in hours worked that is 53.994 hours less than if the individual was not a homeowner.

C\) For individuals who are not homeowners, a one-unit increase in education is associated with a 6.9% increase in hours worked.

D\) A 1% increase in education is associated with an 8.3 hour increase of hours worked.

E\) The sample size is likely smaller because observations in the sample may have worked 0 hours, which would not produce an interperetable value when logged.

### Question 7

The most accurate definition is (B)

### Question 8

\(B\) and (E)

### Question 9

They could use sample weights to adjust the proportions of groups in their sample to reflect the proportions of groups in the larger population.

### Question 10

\(A\)

# Part Two

```{r}
library(tidyverse)
library(modelsummary)
library(fixest)
```

```{r}
d <- read.csv("data/dengue.csv")
```

### Question 1

```{r}
m1 <- lm(NoYes ~ humid + temp,
         data = d)

modelsummary(m1)
```

### Question 2

```{r}
m2 <- glm(NoYes ~ humid + temp,
          family = "binomial",
          data = d)

modelsummary(m2)
```

For each one degree increase in temperature, the odds of dengue increase by 4.1%. For each one unit increase in humidity, the odds of dengue increase by 35.7%.

### Question 3

```{r}
m3 <- lm(humid ~ temp,
         data = d,
         na.action = na.exclude)

modelsummary(m3)
```

```{r}
d$residuals <- residuals(m3)
```

```{r}
ggplot(data = d,
       aes(temp, residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red")
```

```{r}
m4 <- feols(humid ~ temp, 
            data = d, 
            se = "hetero")

modelsummary(m4)
```

```{r}
modelsummary(list(m3, m4))
```

Despite the plot potentially looking like heteroskedasticity, the standard errors did not substantively change, so I would say there is not a meaningful level of heteroskedasticity.

### Question 4

```{r}
m5 <- feols(log(humid) ~ temp,
            data = d,
            se = "hetero")

modelsummary(m5)
```

For a one degree increase in temperature, we would expect to see a 5.8% increase in humidity.

### Question 5

```{r}
ggplot(data = d,
       aes(temp, humid)) +
  geom_point()

ggplot(data = d,
       aes(temp, log(humid))) +
  geom_point()
```

The original graph looks vaguely like an exponential curve, which suggests logging the y variable. The second plot shows that logging the y variable transforms the relationship to more resemble a linear relationship.
