---
title: "Stats_Homework_05"
format: html
---

```{r}
library(tidyverse)
library(lme4)
```

```{r}
d <- readRDS("data/ess5000.rds")
```

# Preliminaries

## 1.

```{r}
psych::describe(d$ppltrst)
var(d$ppltrst)
```

The mean of ppltrst is 4.87, and the variance is 5.55.

## 2.

```{r}
ggplot(d,
       aes(ppltrst)) +
  geom_histogram() +
  theme_minimal()
```

The modal value is 5

## 3.

```{r}
ggplot(d, 
       aes(ppltrst, fill = ppltrst)) +
  geom_bar() +
  facet_wrap(~ cntry) +
  theme_minimal()

ggplot(d, 
       aes(ppltrst, y = cntry)) +
  geom_count(color = "blue") + 
  theme_minimal()

```

Since many of the countries have very little data, and those with a lot of data follow a very similar distribution, I would expect the means to shrink towards the overall mean.

# Single Dimension

## 4.

```{r}
m1 <- lmer(ppltrst ~ (1 | cntry), data = d)

summary(m1)
```

## 5.

```{r}
0.9131/(5.1736+0.9131)
```

## 6.

```{r}
performance::icc(m1)
```

## 7.

This means that 15% of the variation is between groups, meaning most of the variation is within the countries.
